---
title: "AlcHep — Analysis"
date: today
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(tableone)
library(survival)
library(tidycmprsk)
library(ggsurvfit)
library(gtsummary)

source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  quiet = TRUE
))
```

## Analysis Cohort

`master` comes from the shared pipeline. The exposure is `etoh_hepatitis`; the competing-risk outcome is `time_to_death_90days_cmp` / `time_to_death_status_90days_cmp` (0 = censored, 1 = death, 2 = transplant).

```{r cohort}
#| code-fold: true
#| code-summary: "Show cohort filter"
analysis <- master %>%
  filter(!is.na(etoh_hepatitis)) %>%
  mutate(
    etoh_hepatitis = factor(etoh_hepatitis, levels = c(0, 1), labels = c("Non-AH", "AH")),
    event_cmp      = factor(time_to_death_status_90days_cmp,
                            levels = c(0, 1, 2),
                            labels = c("Censored", "Death", "Transplant"))
  )
```

## Table 1 — Demographics by `etoh_hepatitis`

Variable list taken from `Requests/AlcHep Data Request 04162026.docx`.

```{r table1}
#| code-fold: true
#| code-summary: "Show Table 1 code"
numerical_vars <- c(
  "age", "scr_admit", "tbili_admit", "inr_admit", "sodium_admit",
  "albumin_admit", "lactate_admit", "baseline_scr",
  "scr_terli_day0", "bun_terli_day0", "ast_terli_day0", "alt_terli_day0",
  "bili_terli_day0", "alb_terli_day0", "inr_terli_day0", "wbc_terli_day0",
  "hgb_terli_day0"
)

categorical_vars <- c(
  "admit_admitservice", "admit_admitroute", "sex_male", "race", "hispanic",
  "cirrhosis_etiology", "hegrade_admit", "admit_ascites",
  "htn_medical_hx", "dm_medical_hx", "ckd_medical_hx", "cad_medical_hx",
  "nsbb_preadmission", "nsbb_type_preadmission", "diuretics_preadmission",
  "nsaid_preadmission", "ace_arb_preadmission",
  "ascites_previous_comp", "refascites_previous_comp", "he_previous_comp",
  "sbp_previous_comp", "nonsbp_previous_comp", "phtn_bleed_previous_comp",
  "hcc_previous_comp", "tips_previous_comp",
  "he_terli_day0", "location_terli_day0", "terli_route_day0",
  "akin", "ckd_baseline",
  "time_to_death_status_90days_cmp", "hrs_responders_cat_2",
  "HRS_reversal_new", "listed_transplant", "transplant_lt",
  "any_ae", "disposition", "transplant_slkt"
)

numerical_vars   <- intersect(numerical_vars,   names(analysis))
categorical_vars <- intersect(categorical_vars, names(analysis))
all_vars         <- c(numerical_vars, categorical_vars)

analysis <- analysis %>%
  mutate(across(all_of(categorical_vars), as.factor),
         across(all_of(numerical_vars),   as.numeric))

create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = analysis,
  strata   = "etoh_hepatitis",
  output   = "/Users/to909/Desktop/Terlipressin projects/AlcHep/Results/Table1_etoh_hepatitis.csv"
)
```

## Unadjusted Fine-Gray — 90-day mortality (AH vs. non-AH)

```{r fg-unadj}
#| code-fold: true
#| code-summary: "Show unadjusted Fine-Gray"
fg_unadj <- tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, event_cmp) ~ etoh_hepatitis,
  data = analysis
)

fg_unadj %>% tbl_regression(exponentiate = TRUE)

cuminc(Surv(time_to_death_90days_cmp, event_cmp) ~ etoh_hepatitis, data = analysis) %>%
  ggcuminc(outcome = "Death") +
  labs(x = "Days from admission", y = "Cumulative incidence of death") +
  add_confidence_interval() +
  add_risktable()
```

## Adjusted Fine-Gray — MELD 3.0 model

Covariates: `age`, `sex_male`, `hispanic`, `admit_meld_3`, `etoh_hepatitis`.

```{r fg-meld3}
#| code-fold: true
#| code-summary: "Show MELD 3.0 adjusted Fine-Gray"
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, event_cmp) ~
    age + sex_male + hispanic + admit_meld_3 + etoh_hepatitis,
  data = analysis
) %>%
  tbl_regression(exponentiate = TRUE)
```

## Adjusted Fine-Gray — ACLF grade model

Covariates: `age`, `sex_male`, `hispanic`, `aclf_grade`, `etoh_hepatitis`.

```{r fg-aclf}
#| code-fold: true
#| code-summary: "Show ACLF-grade adjusted Fine-Gray"
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, event_cmp) ~
    age + sex_male + hispanic + aclf_grade + etoh_hepatitis,
  data = analysis
) %>%
  tbl_regression(exponentiate = TRUE)
```
