---
title: "AlcHep — Analysis"
date: today
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| include: false
library(tidyverse)
library(readxl)
library(tableone)
library(survival)
library(tidycmprsk)
library(ggsurvfit)
library(gtsummary)
library(riskRegression)
library(prodlim)

source(knitr::purl(
  "/Users/to909/Desktop/Terlipressin projects/Terlipressin/code/shared_pipeline.qmd",
  quiet = TRUE
))
```

```{r helpers}
#| include: false
# Predict adjusted CIF curves for AH vs non-AH from a riskRegression::FGR fit.
# Continuous covariates are held at their median; factor covariates at their
# modal level. The exposure variable is varied across its levels.
adjusted_cif <- function(formula, data, exposure = "etoh_hepatitis",
                         times = 0:90,
                         cause = 1) {           # cause = 1 -> Death (outcome of interest)
                                                # cause = 2 -> Transplant (competing risk)
  fg <- riskRegression::FGR(formula, data = data, cause = cause)

  covars <- setdiff(all.vars(formula)[-c(1, 2)], exposure)
  profile <- lapply(covars, function(v) {
    x <- data[[v]]
    if (is.numeric(x)) {
      median(x, na.rm = TRUE)
    } else {
      lvls <- levels(factor(x))
      modal <- names(sort(table(x), decreasing = TRUE))[1]
      factor(modal, levels = lvls)
    }
  })
  names(profile) <- covars
  profile[[exposure]] <- levels(data[[exposure]])
  newdata <- do.call(expand.grid, c(profile, stringsAsFactors = FALSE))

  risk <- riskRegression::predictRisk(fg, newdata = newdata, times = times)
  colnames(risk) <- times

  as.data.frame(risk) |>
    dplyr::mutate(!!exposure := newdata[[exposure]]) |>
    tidyr::pivot_longer(-dplyr::all_of(exposure),
                        names_to = "time", values_to = "cif") |>
    dplyr::mutate(time = as.numeric(time))
}

plot_adjusted_cif <- function(df, exposure = "etoh_hepatitis",
                              title = NULL, subtitle = NULL) {
  ggplot(df, aes(x = time, y = cif,
                 color = .data[[exposure]], linetype = .data[[exposure]])) +
    geom_step(linewidth = 1) +
    scale_x_continuous(breaks = seq(0, 90, by = 30),
                       limits = c(0, 90),
                       expand = c(0, 0)) +
    scale_y_continuous(limits = c(0, 1), expand = c(0, 0)) +
    labs(x = "Days from admission",
         y = "Adjusted CIF — Death",
         color = "Alcoholic hepatitis",
         linetype = "Alcoholic hepatitis",
         title = title,
         subtitle = subtitle,
         caption = "Outcome: death (cause 1)  |  Competing risk: liver transplant (cause 2)") +
    theme_minimal(base_size = 12) +
    theme(legend.position = "bottom",
          plot.caption = element_text(face = "italic"))
}
```

## Analysis Cohort

`master` comes from the shared pipeline. Throughout this analysis:

- **Exposure:** `etoh_hepatitis` — alcoholic hepatitis (AH) vs. non-AH
- **Primary outcome (event of interest, cause = 1):** death within 90 days
- **Competing risk (cause = 2):** liver transplant within 90 days
- **Censoring (cause = 0):** alive and not transplanted at 90 days / lost to follow-up

Time and status come from the shared pipeline variables `time_to_death_90days_cmp` / `time_to_death_status_90days_cmp` (0 = censored, 1 = death, 2 = transplant).

```{r cohort}
#| code-fold: true
#| code-summary: "Show cohort filter"
# Cause coding for the competing-risk event factor `event_cmp`:
#   level 1 "Censored"   -> censoring        (numeric code 0)
#   level 2 "Death"      -> OUTCOME OF INTEREST (numeric code 1)
#   level 3 "Transplant" -> COMPETING RISK   (numeric code 2)
# This ordering matters: tidycmprsk::crr() treats the first level as the
# censoring code, and `failcode` selects which remaining level is the cause
# of interest (all other non-censoring levels become competing risks).
analysis <- master %>%
  filter(!is.na(etoh_hepatitis)) %>%
  mutate(
    etoh_hepatitis = factor(etoh_hepatitis, levels = c(0, 1), labels = c("Non-AH", "AH")),
    event_cmp      = factor(time_to_death_status_90days_cmp,
                            levels = c(0, 1, 2),
                            labels = c("Censored", "Death", "Transplant"))
  )
```

## Table 1 — Demographics by `etoh_hepatitis`

Variable list taken from `Requests/AlcHep Data Request 04162026.docx`.

```{r table1}
#| code-fold: true
#| code-summary: "Show Table 1 code"
numerical_vars <- c(
  "age", "scr_admit", "tbili_admit", "inr_admit", "sodium_admit",
  "albumin_admit", "lactate_admit", "baseline_scr",
  "scr_terli_day0", "bun_terli_day0", "ast_terli_day0", "alt_terli_day0",
  "bili_terli_day0", "alb_terli_day0", "inr_terli_day0", "wbc_terli_day0",
  "hgb_terli_day0"
)

categorical_vars <- c(
  "admit_admitservice", "admit_admitroute", "sex_male", "race", "hispanic",
  "cirrhosis_etiology", "hegrade_admit", "admit_ascites",
  "htn_medical_hx", "dm_medical_hx", "ckd_medical_hx", "cad_medical_hx",
  "nsbb_preadmission", "nsbb_type_preadmission", "diuretics_preadmission",
  "nsaid_preadmission", "ace_arb_preadmission",
  "ascites_previous_comp", "refascites_previous_comp", "he_previous_comp",
  "sbp_previous_comp", "nonsbp_previous_comp", "phtn_bleed_previous_comp",
  "hcc_previous_comp", "tips_previous_comp",
  "he_terli_day0", "location_terli_day0", "terli_route_day0",
  "akin", "ckd_baseline",
  "time_to_death_status_90days_cmp", "hrs_responders_cat_2",
  "HRS_reversal_new", "listed_transplant", "transplant_lt",
  "any_ae", "disposition", "transplant_slkt"
)

numerical_vars   <- intersect(numerical_vars,   names(analysis))
categorical_vars <- intersect(categorical_vars, names(analysis))
all_vars         <- c(numerical_vars, categorical_vars)

analysis <- analysis %>%
  mutate(across(all_of(categorical_vars), as.factor),
         across(all_of(numerical_vars),   as.numeric))

create_table_one(
  all_vars = all_vars,
  cat_vars = categorical_vars,
  data     = analysis,
  strata   = "etoh_hepatitis",
  output   = "/Users/to909/Desktop/Terlipressin projects/AlcHep/Results/Table1_etoh_hepatitis.csv"
)
```

## Unadjusted Fine-Gray — 90-day mortality (AH vs. non-AH)

**Outcome:** death within 90 days (cause 1) &nbsp;·&nbsp; **Competing risk:** liver transplant (cause 2)

```{r fg-unadj}
#| code-fold: true
#| code-summary: "Show unadjusted Fine-Gray"
# Outcome of interest = "Death"; competing risk = "Transplant"
# (`failcode` is matched against the levels of `event_cmp`).
fg_unadj <- tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, event_cmp) ~ etoh_hepatitis,
  data     = analysis,
  failcode = "Death"   # cause of interest -> all other non-censoring levels are competing risks
)

fg_unadj %>%
  tbl_regression(exponentiate = TRUE) %>%
  modify_caption(
    "Unadjusted subdistribution hazard ratio — Outcome: death (cause 1); Competing risk: liver transplant (cause 2)"
  )

# Aalen–Johansen non-parametric CIF (used for both the plot and the day-90 estimates below)
ci_unadj <- cuminc(
  Surv(time_to_death_90days_cmp, event_cmp) ~ etoh_hepatitis,
  data = analysis
)

ci_unadj %>%
  ggcuminc(outcome = "Death") +
  labs(
    title    = "Unadjusted 90-day cumulative incidence of death",
    subtitle = "AH vs. non-AH",
    x        = "Days from admission",
    y        = "Cumulative incidence — Death",
    caption  = "Outcome: death (cause 1)  |  Competing risk: liver transplant (cause 2)"
  ) +
  add_confidence_interval() +
  add_risktable() +
  scale_x_continuous(breaks = seq(0, 90, by = 30),
                     limits = c(0, 90),
                     expand = c(0, 0)) +
  theme(plot.caption = element_text(face = "italic"))
```

### 90-day cumulative incidence of death (AH vs. non-AH)

Non-parametric **Aalen–Johansen** estimator at *t* = 90 days, accounting for liver transplant as a competing risk. 95% CIs use the log–log transformation.

```{r ci90-table}
#| code-fold: true
#| code-summary: "Show 90-day CIF table"
ci90_raw <- ci_unadj %>%
  tidy(times = 90) %>%
  filter(outcome == "Death")

ci90_raw %>%
  transmute(
    Group              = strata,
    `CIF at 90d (%)`   = sprintf("%.1f", 100 * estimate),
    `95% CI (%)`       = sprintf("%.1f – %.1f", 100 * conf.low, 100 * conf.high)
  ) %>%
  knitr::kable(
    align   = "lrr",
    caption = "Aalen–Johansen 90-day cumulative incidence of death (competing risk: liver transplant)."
  )
```

```{r ci90-sentence}
#| code-fold: true
#| code-summary: "Show inline sentence"
#| results: asis
ah  <- ci90_raw %>% filter(strata == "AH")
nah <- ci90_raw %>% filter(strata == "Non-AH")

cat(glue::glue(
  "At 90 days the cumulative incidence of death accounting for liver transplant ",
  "as a competing risk was **{sprintf('%.1f', 100*ah$estimate)}%** in the AH group ",
  "(95% CI {sprintf('%.1f', 100*ah$conf.low)}–{sprintf('%.1f', 100*ah$conf.high)}) ",
  "and **{sprintf('%.1f', 100*nah$estimate)}%** in the non-AH group ",
  "(95% CI {sprintf('%.1f', 100*nah$conf.low)}–{sprintf('%.1f', 100*nah$conf.high)})."
))
```

## Adjusted Fine-Gray — MELD 3.0 model

**Outcome:** death within 90 days (cause 1) &nbsp;·&nbsp; **Competing risk:** liver transplant (cause 2)
&nbsp;·&nbsp; **Covariates:** `age`, `sex_male`, `hispanic`, `admit_meld_3`, `etoh_hepatitis`.

```{r fg-meld3}
#| code-fold: true
#| code-summary: "Show MELD 3.0 adjusted Fine-Gray"
# Outcome of interest = "Death"; competing risk = "Transplant"
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, event_cmp) ~
    age + sex_male + hispanic + admit_meld_3 + etoh_hepatitis,
  data     = analysis,
  failcode = "Death"   # competing risk: "Transplant"
) %>%
  tbl_regression(exponentiate = TRUE) %>%
  modify_caption(
    "Adjusted (MELD 3.0) subdistribution hazard ratio — Outcome: death (cause 1); Competing risk: liver transplant (cause 2)"
  )
```

```{r fg-meld3-cif}
#| code-fold: true
#| code-summary: "Show adjusted CIF (MELD 3.0 model)"
cif_meld3 <- adjusted_cif(
  prodlim::Hist(time_to_death_90days_cmp, time_to_death_status_90days_cmp) ~
    age + sex_male + hispanic + admit_meld_3 + etoh_hepatitis,
  data = analysis
)

plot_adjusted_cif(
  cif_meld3,
  title    = "Adjusted 90-day CIF of death — MELD 3.0 model",
  subtitle = "Continuous covariates at median, categoricals at modal level"
)
```

## Adjusted Fine-Gray — ACLF grade model

**Outcome:** death within 90 days (cause 1) &nbsp;·&nbsp; **Competing risk:** liver transplant (cause 2)
&nbsp;·&nbsp; **Covariates:** `age`, `sex_male`, `hispanic`, `aclf_grade`, `etoh_hepatitis`.

```{r fg-aclf}
#| code-fold: true
#| code-summary: "Show ACLF-grade adjusted Fine-Gray"
# Outcome of interest = "Death"; competing risk = "Transplant"
tidycmprsk::crr(
  Surv(time_to_death_90days_cmp, event_cmp) ~
    age + sex_male + hispanic + aclf_grade + etoh_hepatitis,
  data     = analysis,
  failcode = "Death"   # competing risk: "Transplant"
) %>%
  tbl_regression(exponentiate = TRUE) %>%
  modify_caption(
    "Adjusted (ACLF grade) subdistribution hazard ratio — Outcome: death (cause 1); Competing risk: liver transplant (cause 2)"
  )
```

```{r fg-aclf-cif}
#| code-fold: true
#| code-summary: "Show adjusted CIF (ACLF model)"
cif_aclf <- adjusted_cif(
  prodlim::Hist(time_to_death_90days_cmp, time_to_death_status_90days_cmp) ~
    age + sex_male + hispanic + aclf_grade + etoh_hepatitis,
  data = analysis
)

plot_adjusted_cif(
  cif_aclf,
  title    = "Adjusted 90-day CIF of death — ACLF-grade model",
  subtitle = "Continuous covariates at median, categoricals at modal level"
)
```
